Upper limits on CN from exocomets transiting $\beta$~Pictoris

### 2025-04-14 src/tex/ms.tex
accepted

\begin{document} 
   \title{Upper limits on CN from exocomets transiting $\beta$~Pictoris}
   \author{M.A. Kenworthy
          \inst{1}
          \and
          E. de Mooij\inst{2}
          \and
          A.\ Brandeker\inst{3}
          \and 
         C. Opitom\inst{4}
          \and
          F. Kiefer\inst{5}
          \and
          A. Fitzsimmons \inst{2}
          }
   \institute{Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
              \email{kenworthy@strw.leidenuniv.nl}
        \and
        Astrophysics Research Centre, School of Mathematics and Physics, Queen’s University Belfast, BT7 1NN Belfast, UK
        \and
        Institutionen f\"{o}r astronomi, Stockholms universitet, AlbaNova universitetscentrum, 106 91, Stockholm, Sweden
        \and
        Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh EH9 3HJ, UK
        \and
        LESIA, Observatoire de Paris, Université PSL, CNRS, 5 Place Jules Janssen, 92190 Meudon, France}

   \date{Received 7 Feb 2025; Accepted XXXX}

@@

\subsection{Initial data processing}
The analysis was started from the 2d extracted spectra (E2DS files).
%
Each order was treated separately, with the main focus on orders 3 and 7 as these orders contain the main \ce{CN} bandhead as well as the \ion{Ca}{ii} H line, which is used to identify the exocomets.
%
All the spectra were transformed into the stellar rest-frame and onto a common wavelength grid using linear interpolation. %For UVES, the \ce{Fe} \hl{NNNN angstrom} line was used to track global wavelength drifts between different exposures which were corrected for at the same time. To measure these, a Gaussian was fit to the line, and the velocity offset from the rest frame position determined.

%
After all the spectra were all transformed into the stellar rest frame, an average nightly spectrum was made from all spectra within a night for each of the nights of observations. This was done after correcting for variations in the effective blaze profile (e.g., due to instrumental changes, differential slit/fiber losses, atmospheric absorption, etc.), as outlined below. Note that after this correction, the %HARPS 
data still have the overall instrumental blaze-profile included.%, while the 2d-spectra from UVES were already corrected for this.

@@

The night of 2018-05-06, which has a high signal-to-noise ratio, was chosen to act as an initial estimate for the blaze function.%(the night of 2018-05-06 and \hl{DATE} for UVES).
%
The individual spectra from this night were averaged together to further boost the signal-to-noise ratio.
%
Subsequently, each individual spectrum was divided by this reference spectrum to highlight the changes in the effective profile.
%
This ratio spectrum was subsequently binned using a bin-width of 81 pixels using a robust algorithm to remove outliers within each bin, clipping points more than 2.7 Median Absolute Deviations (MAD) away from the median of each bin.
%
This binned ratio was then fitted using a second order polynomial.
%
The original spectrum was finally divided by this polynomial to correct for the changes.

@@

After the blaze-variation corrections, all the spectra within a night are combined in order to further boost the signal-to-noise and improve the detection limits.
%
Each individual spectrum was weighted using the average signal-to-noise ratio provided in the header for orders 0 and 11 (3780\,\AA{} - 4075\AA).%\hl{, and OO and PP} for HARPS and UVES respectively.

@@

%For both HARPS and UVES w
We first perform a second iteration of blaze corrections using a stellar spectrum constructed by averaging the spectra from periods with a low comet activity (see Sect.~\ref{sect:FEBid}), similar to the method before, however, now using a wider binsize of 150 pixels and clipping outliers at 4.5 MAD away from the median.
%
%For HARPS
This is done separately pre- and post-fiber change. The correction is done similar to Sect.~\ref{sect:blaze}, however, an 8th order polynomial was used.

@@

After the initial data analysis, we proceed to with our search for lines from \ce{CN}.
%
Although we are interested in lines from the exocomets, we first co-add the data from all the individual epochs, in the stellar rest-frame, to obtain a deep search for circumstellar \ce{CN} absorption.
%
These coadded spectra %for \ac{harps} and \ac{uves} 
are shown in Figure~\ref{fig:spec_CN_stellar_frame}.
%
No obvious feature from the \ce{CN} band jumps out, and we need to combine the signal from the lines within the \ce{CN} band to improve the detection limits.
%
Since the exocomets have a clear radial velocity with respect to \bp{}, and since each exocomet will have its own distinct velocity, we selected the 36 epochs with the strongest exocometary features, a seen in the \ion{Ca}{ii} H line.

@@

The comet features in the \ion{Ca}{ii} H-line ($\lambda=3968$ \AA) were identified after removing the overall stellar line-profile and avoiding the absorption from the circumstellar disk by excluding the central 8 \kms{}.
%
The deepest absorption feature in each observing night was selected, and the location and depth stored.
%
For the comet sample we selected features deeper than 60\% %\hl{and 40\%} for HARPS and UVES, respectively, 
while spectra with features below 15\% %for both HARPS and UVES 
were flagged as `low activity'.
%
The distribution of feature locations and depths is shown in Figure~\ref{fig:FEB_velocity}. The dates, velocities and depths of the selected exocomets can be found in Tab.~\ref{tab:FEB_data}

@@

When creating the CCFs, the region around the narrow, circumstellar \ion{Fe}{I} lines in this wavelength region (see Figs.~\ref{fig:spec_CN_stellar_frame}~and~\ref{fig:spec_CN_comet_frame}) were masked to avoid contamination, especially for the higher temperature CN models.
%
We masked the lines detected in \citep{Kiefer2018}, using a width of $\pm$0.35~$\AA$ for the \ion{Fe}{I} 3859.9$\AA$ line, while for the other lines a width of $\pm$0.15$\AA$ was used.

Even after cross-correlating, the stacked CCFs (see Fig.~\ref{fig:ccf_mean} do not show any clear signals, and we therefore proceed to estimate our detection limits.

@@

The authors thank the Lorentz Centre at Leiden University for organising the workshop ``Exocomets: Understanding the Composition of Planetary Building Blocks'' which led to the study in this paper. 
%
Part of this work was supported by the German \emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project number Ts~17/2--1.

AF acknowledges support from STFC award ST/T00021X/1. EdM acknowledges support from STFC award ST/X00094X/1.
%
We made use of the {\tt Python} programming language \citep{rossum1995} and the open-source {\tt Python} packages {\tt numpy} \citep{walt2011}, {\tt matplotlib} \citep{hunter2007}, {\tt astropy} \citep{astropy2013}.
%      
An online repository with materials used in this work is available at \url{https://github.com/mkenworthy/BetaPicCN}

>>>>

\begin{document} 
   \title{Upper limits on CN from exocomets transiting $\beta$~Pictoris}
   \author{M.A. Kenworthy
          \inst{1}
          \and
          E. de Mooij\inst{2}
          \and
          A.\ Brandeker\inst{3}
          \and 
         C. Opitom\inst{4}
          \and
          F. Kiefer\inst{5}
          \and
          A. Fitzsimmons \inst{2}
          }
   \institute{Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\\
              \email{kenworthy@strw.leidenuniv.nl}
        \and
        Astrophysics Research Centre, School of Mathematics and Physics, Queen's University Belfast, BT7 1NN Belfast, UK
        \and
        Institutionen f\"{o}r astronomi, Stockholms universitet, AlbaNova universitetscentrum, 106 91, Stockholm, Sweden
        \and
        Institute for Astronomy, University of Edinburgh, Royal Observatory, Edinburgh EH9 3HJ, UK
        \and
        LESIA, Observatoire de Paris, Universit\'{e} PSL, CNRS, 5 Place Jules Janssen, 92190 Meudon, France}

   \date{Received 7 Feb 2025; Accepted 13 Apr 2025}

@@

\subsection{Initial data processing}
The analysis was started from the 2d extracted spectra (E2DS files).
%
Each order was treated separately, with the main focus on orders 3 and 7 as these orders contain the main \ce{CN} bandhead as well as the \ion{Ca}{ii} H line, which is used to identify the exocomets.
%
All the spectra were transformed into the stellar rest-frame and onto a common wavelength grid using linear interpolation.

After all the spectra were all transformed into the stellar rest frame, an average nightly spectrum was made from all spectra within a night for each of the nights of observations. This was done after correcting for variations in the effective blaze profile (e.g., due to instrumental changes, differential slit/fiber losses, atmospheric absorption, etc.), as outlined below. Note that after this correction, the data still have the overall instrumental blaze-profile included.

@@

The night of 2018-05-06, which has a high signal-to-noise ratio, was chosen to act as an initial estimate for the blaze function.
%
The individual spectra from this night were averaged together to further boost the signal-to-noise ratio.
%
Subsequently, each individual spectrum was divided by this reference spectrum to highlight the changes in the effective profile.
%
This ratio spectrum was subsequently binned using a bin-width of 81 pixels using a robust algorithm to remove outliers within each bin, clipping points more than 2.7 Median Absolute Deviations (MAD) away from the median of each bin.
%
This binned ratio was then fitted using a second order polynomial.
%
The original spectrum was finally divided by this polynomial to correct for the changes.

@@

After the blaze-variation corrections, all the spectra within a night are combined in order to further boost the signal-to-noise and improve the detection limits.
%
Each individual spectrum was weighted using the average signal-to-noise ratio provided in the header for orders 0 and 11 (3780\,\AA{} - 4075\AA).

@@

We first perform a second iteration of blaze corrections using a stellar spectrum constructed by averaging the spectra from periods with a low comet activity (see Sect.~\ref{sect:FEBid}), similar to the method before, however, now using a wider binsize of 150 pixels and clipping outliers at 4.5 MAD away from the median.
%
This is done separately pre- and post-fiber change. The correction is done similar to Sect.~\ref{sect:blaze}, however, an 8th order polynomial was used.

@@

After the initial data analysis, we proceed to with our search for lines from \ce{CN}.
%
Although we are interested in lines from the exocomets, we first co-add the data from all the individual epochs, in the stellar rest-frame, to obtain a deep search for circumstellar \ce{CN} absorption.
%
These coadded spectra are shown in Figure~\ref{fig:spec_CN_stellar_frame}.
%
No obvious feature from the \ce{CN} band jumps out, and we need to combine the signal from the lines within the \ce{CN} band to improve the detection limits.
%
Since the exocomets have a clear radial velocity with respect to \bp{}, and since each exocomet will have its own distinct velocity, we selected the 36 epochs with the strongest exocometary features, a seen in the \ion{Ca}{ii} H line.

@@

The comet features in the \ion{Ca}{ii} H-line ($\lambda=3968$ \AA) were identified after removing the overall stellar line-profile and avoiding the absorption from the circumstellar disk by excluding the central 8 \kms{}.
%
The deepest absorption feature in each observing night was selected, and the location and depth stored.
%
For the comet sample we selected features deeper than 60\% while spectra with features below 15\% were flagged as `low activity'.
%
The distribution of feature locations and depths is shown in Figure~\ref{fig:FEB_velocity}. The dates, velocities and depths of the selected exocomets can be found in Tab.~\ref{tab:FEB_data}

@@

When creating the CCFs, the region around the narrow, circumstellar \ion{Fe}{I} lines in this wavelength region (see Figs.~\ref{fig:spec_CN_stellar_frame}~and~\ref{fig:spec_CN_comet_frame}) were masked to avoid contamination, especially for the higher temperature CN models.
%
We masked the lines detected in \citep{Kiefer2018}, using a width of $\pm0.35$\AA{} for the \ion{Fe}{I} 3859.9\AA{} line, while for the other lines a width of $\pm0.15$\AA{} was used.

Even after cross-correlating, the stacked CCFs (see Fig.~\ref{fig:ccf_mean}) do not show any clear signals, and we therefore proceed to estimate our detection limits.

@@

The authors thank the Lorentz Centre at Leiden University for organising the workshop ``Exocomets: Understanding the Composition of Planetary Building Blocks'' which led to the study in this paper. 
%
Part of this work was supported by the German \emph{Deut\-sche For\-schungs\-ge\-mein\-schaft, DFG\/} project number Ts~17/2--1.
%
AF acknowledges support from STFC award ST/T00021X/1. EdM acknowledges support from STFC award ST/X00094X/1.
%
We made use of the {\tt Python} programming language \citep{rossum1995} and the open-source {\tt Python} packages {\tt numpy} \citep{walt2011}, {\tt matplotlib} \citep{hunter2007}, {\tt astropy} \citep{astropy2013}.
%      
An online repository with materials used in this work is available at \url{https://github.com/mkenworthy/BetaPicCN}
